Thesis Title

### 2022-02-20 chapters/accronims.tex
update

Global Burden of Disease (GBD)
World Health Organization (WHO)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)
T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)
PubMed Central\textsuperscript{\small\textregistered} (PMC)
Public health surveillance systems (PHSS)

>>>>

Global Burden of Disease (GBD)
World Health Organization (WHO)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)
T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)
PubMed Central\textsuperscript{\small\textregistered} (PMC)
Public health surveillance systems (PHSS)
dideoxynucleotide triphosphate (ddNTP) 
deoxynucleotide triphosphate (dNTP)
single molecule real-time sequencing (SMRT)
Whole genome sequencing (WGS)

### 2022-02-20 chapters/introduction.tex
update

\section{The global impact of microbial pathogens}

@@

\subsection{Current standards for diagnostic in clinical microbiology}

@@

\subsubsection{Bacterial infections}

@@

\subsubsection{Viral infections}

@@

\subsubsection{Fungal infections}

@@

\subsubsection{Parasitic infections}

@@

\subsection{Surveillance and infection prevention in public health}

@@

\section{A genomic approach to clinical microbiology}

@@

\subsection{Twenty five years of microbial genome sequencing}

Since the  the discovery of the structure of DNA \citep{watson_molecular_1953}, great strides have been made in understanding the complexity and diversity of genomes in health and disease. The ability to determine the sequence in a DNA molecule

The development and commercialisation of high-throughput, massively parallel sequencing, has democratised sequencing by offering individual laboratories, either in research or in health, access to the technology.

\subsubsection{The first generation of DNA sequencing}

@@

\subsubsection{The second generation of DNA sequencing}

@@

\subsubsection{Sequencing by hybridisation}

@@

\subsubsection{Sequencing by synthesis}

@@

\subsubsection{The third generation of DNA sequencing}

@@

\subsection{Metagenomics}

>>>>

\section{The global impact of microbial pathogens} \label{sec:global_impact}

@@

\subsection{Current standards for diagnostic in clinical microbiology} \label{ssec:current_standards}

@@

\subsubsection{Bacterial infections} \label{sssec:bacterial}

@@

\subsubsection{Viral infections} \label{sssec:viral}

@@

\subsubsection{Fungal infections} \label{sssec:fungal}

@@

\subsubsection{Parasitic infections} \label{sssec:parasites}

@@

\subsection{Surveillance and infection prevention in public health} \label{ssec:survaillance}

@@

\section{A genomic approach to clinical microbiology} \label{sec:genomics_approach}

@@

\subsection{Twenty five years of microbial genome sequencing} \label{ssec:sequencing}

Since the  the discovery of the structure of DNA \citep{watson_molecular_1953}, great strides have been made in understanding the complexity and diversity of genomes in health and disease. The ability to determine the sequence in a DNA molecule  
The development and commercialisation of high-throughput, massively parallel sequencing, has democratised sequencing by offering individual laboratories, either in research or in health, access to the technology.

\subsubsection{The first generation of DNA sequencing} \label{ssec:1st_gen_seq}

@@

\subsubsection{The second generation of DNA sequencing} \label{ssec:2nd_gen_seq}

@@

\paragraph{Sequencing by hybridisation} \label{sssec:2nd_gen_seq_hybrid} \mbox{}\\

@@

\paragraph{Sequencing by synthesis} \label{sssec:2nd_gen_seq_synth} \mbox{}\\

@@

\subsubsection{The third generation of DNA sequencing} \label{ssec:3rd_gen_seq}

@@

\subsection{DNA sequencing in clinical diagnosis and surveillance} \label{ssec:sequencing_diagnosis}

Whole genome sequencing (WGS) is becoming one of the most widely used applications of microbial genome sequencing. The major advantage of WGS is to yield all the available DNA information content on isolates in a single rapid step following culture (sequencing without culture will be discussed in the \secref{ssec:metagenomics}). In principle, after obtaining a pure culture, either bacterial (see \secref{sssec:bacterial}), viral (see \secref{sssec:viral}) of fungal (see \secref{sssec:fungal}), the data from sequencing contain all the information currently used for diagnostic and typing needs, and much more, thus opening the prospect for large-scale research into pathogen genotype-phenotype associations from routinely collected data \citep{didelot_transforming_2012}.
The cost of producing massive amounts of information requires a new framework with expert handling and processing of computer-driven genomic information, as well as capable computational infrastructures, but through this technology, researchers and clinicians can obtain the most comprehensive view of genomic information and associated biological implications, transforming clinical diagnostics and the detection and surveillance of outbreaks. \citep{cirulli_uncovering_2010, noauthor_genomic_2019, goodwin_coming_2016}.

Targeted sequencing is also proving invaluable to clinical microbial and research, not only by allowing more individual samples to be sequenced within a single run, significantly reducing costs and the amount of data generated while, but also, due to the smaller target size, obtaining results with very high confidence due to the high coverage obtained \citep{goodwin_coming_2016}. This has been particularly useful in viral genomics where sections, such as the capsid, or the complete viral genome can be selectively targeted directly from the suspected sample, offering a more time-effective method to achieve the same output as traditional nucleic acid amplification methods \citep{cassedy_virus_2021}.

WGS has be used in the routine laboratory workflow when typing of pathogens by a method having the highest possible discriminatory power is required, either through single nucleotide polymorphism (SNP) or core-genome/whole genome MLST analysis, for example during hospital outbreaks \citep{tagini_bacterial_2017}. Additionally, in bacterial diagnostics, WGS can be used to reveal the presence of AMR genes, or genes associated with virulence and pathogenicity, as well as to discover new genetic mechanisms for the three previously defined important clinical features of a bacterium \citep{rossen_practical_2018}. The implementation of WGS in routine diagnostics requires several adaptations in the laboratory workflow, from the ‘wet’ laboratory part (extraction, library preparation, sequencing), to the ‘dry’ bioinformatics part where genomic data is analysed and its results interpreted by specialised personnel \citep{rossen_practical_2018}.

Currently, sequencing technologies are used in a case-by-case approach, with its adoption being much more present in a research setting than in a diagnostic one. Sequencing is mostly used after a diagnostic through the identification of the causative agent has already been performed. Although substantial advances have been made in reducing response time, most of the current systems do not yet generate enough data fast enough for a truly rapid response for it to be used in the clinical setting \citep{goodwin_coming_2016}. High-throughput DNA sequencing has found additional new applications in drug discovery and in functional genomics with, for example, SNP-based analysis to identify new drug targets \citep{loman_twenty_2015}.

Although the second-generation DNA sequencing methods have shed light on fundamental aspects of microbial ecology and function, they suffer from issues associated with short read length (see \ref{ssec:2nd_gen_seq}) and cannot reliably reconstruct long repeats because of uncertainties in mapping read, even when paired end sequencing is used. Third generation sequencing methods (see \ref{ssec:3rd_gen_seq}) have become increasingly used in microbiology, although their accuracy and low throughput make it challenging to implement in a clinical diagnosis setting. %CITATION NEEDED!!!

Most notably, WGS has become a common tool in surveillance and infection prevention, allowing for pathogen identification and tracking, establishing transmission routes and outbreak control \citep{lo_genomics_2020}. In bacterial infections, initiatives such as Pathogenwatch\footnote{\url{https://pathogen.watch/}} offers a web-based platform for AMR analysis and phylogeny generation of \textit{Campylobacter}, \textit{Klebsiella}, \textit{Neisseria gonorrhoeae}, \textit{Staphylococcus aureus}, and \textit{Salmonella Typhi} \citep{afolayan_overcoming_2021}. The Center for Genomic Epidemiology website\footnote{\url{https://www.genomicepidemiology.org/}} offers services for phylogenetic tree building and AMR prediction. Chewie Nomenclature Server\footnote{\url{https://chewbbaca.online/}} allows users to share genome-based gene-by-gene typing schemas and to maintain a common nomenclature, simplifying the comparison of results \citep{mamede_chewie_2021}. Enterobase\footnote{\url{https://enterobase.warwick.ac.uk/}} allows for the analysis and visualisation of genomic variation within enteric bacteria \citep{zhou_enterobase_2020}. Microreact\footnote{\url{https://microreact.org/}}, from the same developers as Pathogenwatch, combines clustering, geographical and temporal data into an interactive visualisation with trees, maps, timelines and tables for a multitude of microorganisms, both bacterial and viral \citep{argimon_microreact_nodate}. Particularly for viruses, GISAID\footnote{\url{https://www.gisaid.org/}}  promotes the rapid sharing of data from all influenza viruses and the coronavirus causing COVID-19, including the genetic sequences and related clinical and epidemiological data \citep{shu_gisaid_2017}. ViPR\footnote{\url{https://www.viprbrc.org/}} provides access to sequence records, gene and protein annotations, immune epitopes, 3D structures, host factor data, and other data types for over 14 viral families, including \textit{Coronaviridae}, from which SARS-CoV-2 belongs to, and \textit{Faviviridae}, the family of Dengue and Zika virus \citep{pickett_virus_2012}. INSaFLU\footnote{\url{https://insaflu.insa.pt/}} supplies public health laboratories and influenza researchers with an web-based suit for effective and timely influenza and SARS-CoV-2 laboratory surveillance, identifying type and subtype/lineage, detection of putative mixed infections and intra-host minor variants \citep{borges_insaflu_2018}. Nextrain\footnote{\url{https://nextstrain.org/}} provide a continually-updated view of publicly available data alongside powerful analytic and visualisation tools o aid epidemiological understanding and improve outbreak response for 10 pathogens: Influenza, SARS-CoV-2, West Nile virus, Mumps, Zika, West African Ebola, Dengue, Measles, Enterovirus D68 and Tuberculosis \citep{hadfield_nextstrain_2018}

\subsection{From genomics to metagenomics} \label{ssec:metagenomics}

loman
l with the relentless rise of the microbiome across the scientific agenda
and in the public eye91, high-throughput sequencing has been harnessed for culture- independent approaches to microbial ecol- ogy and even for diagnosis.

molecular bar-coding approaches could be combined with high-throughput sequencing to achieve unprecedented depths of coverage in microbial community profiling

 In recent years, there has been a growing interest in using metagenomics to deliver a new culture-independent para- digm in diagnostic microbiology,

leo
one existing challenge in diagnostics remains; a priori knowledge of what to expect from a particular clinical sample or patient. In most cases, a priori knowledge is enough to request the most appropriate test, such as multiplexed panels or specific culture media, but this is not always the case.
